ABC Group Level Fits

### 2024-01-06 Model/abc_group.qmd
abc visuals

1. Load the group level ABC results - i.e. the posterior distribution for the c and lr parameters. 
2. Preview the posterior distribution.

```{r}
pacman::p_load(tidyverse, data.table,dtplyr, here, patchwork, conflicted)
conflict_prefer_all("dplyr", quiet = TRUE)

group_prior=abc_1M_p001 = readRDS(here::here("data/abc_1M_rmse_p001.rds"))
group_prior=abc_2M_p001 = readRDS(here::here("data/abc_2M_rmse_p0001.rds"))
names(group_prior)
str(group_prior$abc_ev[[1]])

tetr_v <- group_prior$abc_ev$teter_results

@@

teter_v_test <- teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test",condit=="Varied")
```

```{r fig.width=12, fig.height=10}

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test") |>
  mutate(facet_label = paste0("rank: ", rank, "\n", "c: ", round(c, 4), "\n", "lr: ", round(lr, 4), "\n", "distance: ", round(distance, 1))) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(~Model)

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test",rank<=10) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(~Model)

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test",rank<=10) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(rank~Model)

>>>>

1. Load the group level ABC results - i.e. the posterior distribution for the c and lr parameters. 
2. Examine the posterior distribution of the c and lr parameters. 
3. Examine the posterior predictive distributions of the test data, and compare to the observed test data.

```{r}
pacman::p_load(tidyverse, data.table, here, patchwork, conflicted)
conflict_prefer_all("dplyr", quiet = TRUE)

walk(c("fun_alm","fun_model", "Display_Functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
dsv <- ds |> filter(condit=="Varied")  
dsc <- ds |> filter(condit=="Constant") 
ind_ds <- ds |> 
  filter(expMode2 == "Test") |> 
  group_by(id, condit, x, expMode2) |> 
  summarise(y = mean(y), .groups = "keep")

tMax=84
avg_dsv <- ds |> filter(condit=="Varied",expMode2=="Train") |> group_by(tr) %>%
  mutate(bandInt2 = sample(rep(c(800, 1000, 1200), each = tMax / 3), tMax, replace = FALSE)[tr]) %>%
  filter(bandInt2 == x) |> select(-bandInt2) |> group_by(tr,condit,x,expMode2) |> summarise(y=mean(y),.groups="keep") |>
  rbind(dsv |> filter(expMode2=="Test") |> group_by(condit,x,expMode2) |> summarise(y=mean(y),tr=1,.groups="keep") ) |> setDT()

avg_dsc <- ds |> filter(condit=="Constant",expMode2=="Train",tr<=tMax) |> group_by(tr, condit,x,expMode2) |> 
  summarise(y=mean(y),.groups="keep") |> 
  rbind(dsc |> filter(expMode2=="Test") |> 
          group_by(condit,x,expMode2) |> 
          summarise(y=mean(y),tr=1,.groups="keep") ) |> 
  setDT()

input_layer =  c(100,350,600,800,1000,1200)
output_layer = input_layer

datasets <- list(
  abc_1M_p001 = readRDS(here::here("data/abc_1M_rmse_p001.rds")),
  abc_1M_p0001 = readRDS(here::here("data/abc_1M_rmse_p0001.rds")),
  abc_2M_p001 = readRDS(here::here("data/abc_2M_rmse_p001.rds")),
  abc_2M_p0001 = readRDS(here::here("data/abc_2M_rmse_p0001.rds")),
   abc_1p5M_p001 = readRDS(here::here("data/abc_1p5M_rmse_p001.rds")),
  abc_1p5M_p0001 = readRDS(here::here("data/abc_1p5M_rmse_p0001.rds"))
)

# group_prior=abc_1M_p001 = readRDS(here::here("data/abc_1M_rmse_p001.rds"))
group_prior=abc_2M_p001 = readRDS(here::here("data/abc_2M_rmse_p0001.rds"))

names(group_prior)
#str(group_prior$abc_ev[[1]])

#tetr_v <- group_prior$abc_ev$teter_results

@@

#teter_v_test <- teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test",condit=="Varied")

```{r fig.width=11, fig.height=8}

get_order_pattern <- function(df, pred_col = "pred",band_tol=50) {

  df <- df[order(df$x), ]
  # Calculate the standard deviation and determine the pattern
  pred_sd <- sd(df[[pred_col]]) / 2.8
  pattern <- ""
  # Iterate through the rows to build the pattern
  for (i in 1:(nrow(df) - 1)) {
    if (abs(df[[pred_col]][i] - df[[pred_col]][i + 1]) <= band_tol) {
      pattern <- paste0(pattern, "=")
    } else if (df[[pred_col]][i] < df[[pred_col]][i + 1]) {
      pattern <- paste0(pattern, "<")
    } else {
      pattern <- paste0(pattern, ">")
    }
  }

  return(pattern)
}

# Apply the function to the dataset
ind_pattern <- ind_ds |> 
  group_by(id, condit, expMode2) |>
  do(order_pattern = get_order_pattern(., pred_col = "y",band_tol=100))

ind_ds <- ind_ds |>
  left_join(ind_pattern, by=join_by(id,condit,expMode2)) |> 
  mutate(order_pattern = unlist(order_pattern))

ord_pattern <- rbind(teter, te, tr)  |> select(sim_dat) |> unnest(sim_dat)  |> 
  filter(expMode2=="Test") |> 
  group_by(condit,rank,Model,Fit_Method) %>% 
  do(order_pattern = unlist(get_order_pattern(.)))

combo_pred <- rbind(teter, te, tr)  |> 
  select(sim_dat) |> 
  unnest(sim_dat)  |> 
  filter(expMode2=="Test") |>  
  left_join(ord_pattern, by=join_by(condit,rank,Model,Fit_Method)) |>
  mutate(order_pattern = unlist(order_pattern))

head(combo_pred)

length(unique(combo_pred$order_pattern))

pattern_freq <- combo_pred %>%
  count(order_pattern) %>%
  arrange(desc(n))

# Print the frequency table
print(pattern_freq)

# Plot the frequency distribution
ggplot(pattern_freq, aes(x = order_pattern, y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Frequency Distribution of Order Patterns",
       x = "Order Pattern",
       y = "Frequency")

combo_pred %>% group_by(condit,Model,Fit_Method) |> count(order_pattern) |> arrange(desc(n)) |> ggplot(aes(x = order_pattern, y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Frequency Distribution of Order Patterns",
       x = "Order Pattern",
       y = "Frequency") +
  ggh4x::facet_nested_wrap(condit~Model)

  combo_pred %>%
  group_by(Model, Fit_Method) %>%
  mutate(n_pattern = n_distinct(order_pattern)) %>%
  group_by(Model, Fit_Method, order_pattern) %>%
  mutate(pat_count = n()) %>%
  ungroup() %>%
  group_by(Model, Fit_Method) %>%
  arrange(Model, Fit_Method, desc(pat_count), order_pattern) %>%
  mutate(pat_rank = dense_rank(desc(pat_count))) |> 
  group_by(Model, Fit_Method, order_pattern) %>%
  # distinct(Model, Fit_Method, order_pattern, .keep_all = TRUE) %>%
  # arrange(Model, Fit_Method, pat_rank)
  filter(pat_rank<=5) |>
  filter(condit=="Varied")  |> 
 ggplot(aes(x=c,y=lr,color=as.factor(order_pattern))) + geom_point() + facet_wrap(Fit_Method~Model,scales="free") 
  # use viridis color palette
 # scale_color_viridis_d(option = "magma", direction = -1)

ordered_labels <- ind_ds |>
  group_by(order_pattern, x,condit) |>
  summarize(y = mean(y), n_id = n_distinct(id)) |>
  arrange(-n_id, order_pattern) |>
  mutate(factor_label = paste0(order_pattern, " (n=", n_id, ")")) %>%
  pull(factor_label) %>%
  unique()

k= ind_ds |>
  group_by(order_pattern, x,condit) |>
  summarize(y = mean(y), n_id = n_distinct(id)) |>
  mutate(n_id = factor(n_id, levels = lorder)) |>
  mutate(factor_label = paste0(order_pattern, " (n=", n_id, ")"),
         factor_label = factor(factor_label, levels = ordered_labels))

k |> filter(condit=="Varied") |>
  ggplot(aes(x = x, y = y,fill=condit),fill=wes_palette("Darjeeling1")[1]) +
  stat_summary(fun = mean, geom = "bar", position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.8), width = 0.25) +
  facet_wrap(~factor_label, scales = "free_x")

k |> filter(condit=="Constant") |>
  ggplot(aes(x = x, y = y),fill=wes_palette("Darjeeling1")[2]) +
  stat_summary(fun = mean, geom = "bar", position = position_dodge(width = 0.8)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.8), width = 0.25) +
  facet_wrap(~factor_label, scales = "free_x")

# Posterior Distribution of Model Parameters

```{r fig.width=12, fig.height=10}
#| eval: true

# Define function to process dataset
process_dataset <- function(name, dataset) {
  sample_size <- str_extract(name, "\\d+p?\\d*?M")
  posterior_cutoff <- str_extract(name, "p\\d+$")

  teter_combined <- map_dfr(dataset, "teter_results", .id = "run_name")
  te_combined <- map_dfr(dataset, "te_results", .id = "run_name")
  tr_combined <- map_dfr(dataset, "tr_results", .id = "run_name")

  plot_title_suffix <- sprintf("%s - %s", sample_size, posterior_cutoff)

  return(list(teter_combined = teter_combined,
              te_combined = te_combined,
              tr_combined = tr_combined,
              plot_title_suffix = plot_title_suffix))
}

create_density_plots <- function(processed_data) {
  teter_combined <- processed_data$teter_combined
  te_combined <- processed_data$te_combined
  tr_combined <- processed_data$tr_combined
  plot_title_suffix <- processed_data$plot_title_suffix

  # Base plot for density
  base_plot <- function(data, x_var, title, show_legend = FALSE) {
    ggplot(data, aes_string(x=x_var, color="Model")) +
      geom_density() +
      facet_wrap(~Group, scales = "free") +
      ggtitle(title) +
      theme(legend.position=ifelse(show_legend, "top", "none"))  # Toggle legend display
  }

  # Create plots, but only show legend on the first plot
  plot1 <- base_plot(teter_combined, "c", sprintf("c posterior - Test & Train - %s", plot_title_suffix),TRUE )
  plot2 <- base_plot(teter_combined, "lr", sprintf("lr posterior - Test & Train - %s", plot_title_suffix))
  plot3 <- base_plot(te_combined, "c", sprintf("c posterior - Test Only - %s", plot_title_suffix))
  plot4 <- base_plot(te_combined, "lr", sprintf("lr posterior - Test Only - %s", plot_title_suffix))
  plot5 <- base_plot(tr_combined, "c", sprintf("c posterior - Train Only - %s", plot_title_suffix))
  plot6 <- base_plot(tr_combined, "lr", sprintf("lr posterior - Train Only - %s", plot_title_suffix))

  # Combine plots and place the one with the legend (plot1) at the end so that the legend is not cut off
  density_combined <- plot1 + plot2 + plot3 + plot4 + plot5 + plot6 + 
    plot_layout(ncol = 2) 
  return(density_combined)
}

create_distance_plots <- function(processed_data) {
  teter_combined <- processed_data$teter_combined
  te_combined <- processed_data$te_combined
  tr_combined <- processed_data$tr_combined
  plot_title_suffix <- processed_data$plot_title_suffix

  # Base plot for distance
  base_dist_plot <- function(data, title, show_legend = FALSE) {
    ggplot(data, aes(x=Group, y=distance, fill=Model)) +
      stat_summary(fun=mean, geom="bar", position=position_dodge(width = 0.8)) +
      stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width = 0.8), width=0.25) +
      ggtitle(title) +
      theme(legend.position=ifelse(show_legend, "top", "none")) # Control the display of the legend
  }

  # Create distance plots
  dist_plot1 <- base_dist_plot(teter_combined, sprintf("Test & Train - %s", plot_title_suffix),TRUE)
  dist_plot2 <- base_dist_plot(te_combined, sprintf("Test Only - %s", plot_title_suffix),)
  dist_plot3 <- base_dist_plot(tr_combined, sprintf("Train Only - %s", plot_title_suffix))

  # Combine plots and place the one with the legend (dist_plot1) at the end
  distance_combined <- dist_plot1 + dist_plot2 + dist_plot3 + 
    plot_layout(ncol = 2) 
  return(distance_combined)
}

# Define function to save plots
save_plots <- function(density_plot, distance_plot, name) {
  sample_size <- str_extract(name, "\\d+p?\\d*?M")
  posterior_cutoff <- str_extract(name, "p\\d+$")

  density_filename <- paste0("assets/tmp_plots/density_plots_combo_abc_", sample_size, "_rmse_", posterior_cutoff, ".png")
  distance_filename <- paste0("assets/tmp_plots/distance_plots_combo_abc_", sample_size, "_rmse_", posterior_cutoff, ".png")

  ggsave(filename = here::here(density_filename), plot = density_plot, width = 10, height = 6)
  ggsave(filename = here::here(distance_filename), plot = distance_plot, width = 10, height = 6)
}

# Process datasets
processed_datasets <- lapply(names(datasets), function(name) process_dataset(name, datasets[[name]]))

density_plots <- invisible(lapply(processed_datasets, create_density_plots))
distance_plots <- lapply(processed_datasets, create_distance_plots)

# Save plots
#invisible(lapply(seq_along(names(datasets)), function(i) save_plots(density_plots[[i]], distance_plots[[i]], names(datasets)[i])))

```{r fig.width=12, fig.height=10}
density_plots
```

```{r fig.width=12, fig.height=10}
distance_plots
```

```{r fig.height=14,fig.width=11}

pat_combo1 <- combo_pred |> group_by(condit,Model,Fit_Method,order_pattern) |> slice_min(rank,n=1)

pat_combo1 |> filter(Model=="EXAM",condit=="Varied") |> ggplot(aes(x = x, y = pred, fill=as.factor(x))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  #ggh4x::facet_grid2(vars(Fit_Method),vars(order_pattern),axes="all")
  #ggh4x::facet_wrap2(vars(Fit_Method,order_pattern), trim_blank=FALSE)
  #ggh4x::facet_nested(~Fit_Method+order_pattern)
  #ggh4x::facet_nested(order_pattern~Fit_Method, scales="free")
  #facet_nested_wrap(vars(Fit_Method,order_pattern), scales="free",trim_blank=FALSE)
  ggh4x::facet_grid2(vars(order_pattern),vars(Fit_Method),render_empty=FALSE)

{pat_combo1 |> filter(condit=="Varied",Fit_Method=="Test & Train",Model=="ALM") |> ggplot(aes(x = x, y = pred, fill=as.factor(x))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_nested_wrap(Model~order_pattern, scales="free",trim_blank=FALSE,ncol=5) } /
pat_combo1 |> filter(condit=="Varied",Fit_Method=="Test & Train",Model=="EXAM") |> ggplot(aes(x = x, y = pred, fill=as.factor(x))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_nested_wrap(Model~order_pattern, scales="free",trim_blank=FALSE,ncol=5) + 
  theme(legend.position="none")

combo_pred |> group_by(condit,order_pattern,Model) |> filter(Fit_Method=="Test & Train",condit=="Varied",Model!="Alt_EXAM")  |> 
  arrange(distance,Model) |>
  slice_max(distance,n=1) |> 
  group_by(Model,x) |> 
  slice_head(n=10) |> 
  mutate(facet_label = paste0(Model,"\n",order_pattern,"\n" ,"c: ", round(c, 4), "; ", "lr: ", round(lr, 2), "; ", "dist: ", round(distance, 0))) |>
  ggplot(aes(x = x, y = pred, fill=as.factor(x))) +
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  stat_halfeye(aes(x=x,y=y),position=position_dodge(), alpha=.6) +
  facet_wrap(~facet_label,ncol=5)

# 
# pat_combo1 |> filter(Model=="EXAM",condit=="Varied", order_pattern=="<<<<=") |> select(Model,Fit_Method,condit,x,pred,y) |> head(20) |> DT::datatable()
# pat_combo1 |> filter(Model=="EXAM",condit=="Varied", order_pattern=="<<<<>") |> select(Model,Fit_Method,condit,x,pred,y) |> head(20) |> DT::datatable()

# Posterior Predictive

## Average of posterior predictive distributions

```{r fig.width=12, fig.height=10}
combo_pred |> select(sim_dat) |> unnest(sim_dat)  |>  filter(expMode2=="Test") |>
  mutate(facet_label = paste0("rank: ", rank, "\n", "c: ", round(c, 4), "\n", "lr: ", round(lr, 4), "\n", "distance: ", round(distance, 1))) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge(), alpha=.75) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  #stat_summary(fun=mean, geom="point", aes(x=x,y=y,color=condit), position=position_dodge()) +
  stat_halfeye(aes(x=x,y=y,color=condit),position=position_dodge()) +
  ggh4x::facet_nested_wrap(Fit_Method~Model) + labs(title = "Average of posterior predictive distributions")

combo_pred |> select(sim_dat) |> unnest(sim_dat)  |> filter(expMode2=="Test",rank<=5) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge(), alpha=.75) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
   stat_halfeye(aes(x=x,y=y,color=condit),position=position_dodge()) +
  facet_wrap(Fit_Method~Model) + labs(caption = "Average of 5 best posterior predictive distributions")

combo_pred |> select(sim_dat) |> unnest(sim_dat)  |> 
  mutate(n_rank=max(rank)) |>
  filter(expMode2=="Test",rank>=n_rank-5) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge(), alpha=.75) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
   stat_halfeye(aes(x=x,y=y,color=condit),position=position_dodge()) +
  facet_wrap(Fit_Method~Model) + labs(caption = "Average of 5 worst posterior predictive distributions")
```

## Testing Patterns - best fits
```{r fig.width=12, fig.height=15}

combo_pred |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test",rank<=4) |>
  ggplot(aes(x = x, y = pred, fill=as.factor(x))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  ggh4x::facet_nested_wrap(condit~Model~Fit_Method~rank, nrow=6)

## Testing Patterns - worst fits
```{r fig.width=12, fig.height=15}

combo_pred |> select(sim_dat) |> unnest(sim_dat) |> mutate(n_rank=max(rank)) |>
  filter(expMode2=="Test",rank>n_rank-4) |>
  ggplot(aes(x = x, y = pred, fill=as.factor(x))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  ggh4x::facet_nested_wrap(condit~Model~Fit_Method~rank, nrow=6)

## Testing Patterns - Average of models
```{r fig.width=12, fig.height=18}

combo_pred |> select(sim_dat) |> unnest(sim_dat) |> mutate(n_rank=max(rank)) |>
  filter(expMode2=="Test",
        # condit=="Varied",
         Model != "Alt_EXAM",
         Fit_Method=="Test & Train", 
        rank>n_rank-30) |>
  ggplot(aes(x = x, y = pred, fill=as.factor(x))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  ggh4x::facet_nested_wrap(condit~Model~rank, nrow=12)
